---
title: "Practices on Lists"
author: "Salah"
date: "11/10/2021"
output: html_document
---
```{r}
my.list <-  list( stud.id =12456 , stud.name =" Ali", 
stud.marks=c(56 ,65 ,68 ,78))
my.list$stud.id
my.list$stud.name
my.list$stud.marks
new.list <- list( age =17 ,    sex=   " Male ")
updated.list <- list(my.list,new.list)
class(updated.list)

```

```{r}
my.list <-  list( stud.id =12456 , stud.name =" Ali", 
stud.marks=c(56 ,65 ,68 ,78))
my.list$stud.id
my.list$stud.name
my.list$stud.marks
new.list <- list( age =17 ,    sex=   " Male ")
updated.list <- list(my.list,new.list)
class(updated.list)
```
# Useing names(..) to modify the attribute names.
```{r}
list    <-   list( ID =154 ,    NAME =" Nasser")
names(list)
names(list) <-  c(" Identification ", " First Name ")
list
```
### Remove list members either by -ve or NULL
```{r}
length(list)
list <- list[-2]
list
```

# Useing names(..) to modify the attribute names.
```{r}
list    <-   list( ID =154 ,    NAME =" Nasser")
names(list)
names(list) <-  c(" Identification ", " First Name ")
list
```
### Remove list members either by -ve or NULL
```{r}
length(list)
list <- list[-2]
list
```
### Creating a list
```{r}
x <- list(a = 2.5,b = TRUE, c = 1:3)
class(x)
typeof(x)
str(x)
```
### Combining two lists 
```{r}
n <- c(2,3,5)
s <- c("aa","bb","cc","dd")
list <- list("First" = s,"Second" =n,"Third" = 3)
list
str(list) #Metadata about the list
```

```{r}
#List slicing 
list[2]
list[c(1,3)]
```
```{r}
list["First"] #Slicing by the name of vectors 
```

### Similar manner to vectors. Integer, logical or character vectors can be used for  indexing.

### Indexing with [ as shown above will give us sublist not the content  inside the component. To retrieve the content, we need to use [[.

```{r}
list["First"] # Single [ returns a list
list[["First"]] # Double [[ returns a vector
typeof(list["First"] ) # List
typeof(list[["First"]]) #Character
```
### We can use $ instead of [[]]
```{r}
list$First 
list$Second
typeof(list$First) #Character 
class(list$First) #Character
typeof(list$Second) #double 
class(list$Second) #Numeric
```
### Further indexing can be accomplished also !
```{r}
list$First[2] 
# Modification as follows,
list$Second <- "New String insead of 2 3 5 Vector"
list
```
### We can delete a component by assigning NULL to it.
```{r}
list[["First"]] <- NULL
list
str(list)
```
