---
title: "Amazon Web Scraping with R"
author: "Salah"
date: "11/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##### The scraped product is "Samsung DC-E Series Commercial LED Displays 32-Inch Screen LED-Lit Monitor (DC32E)" from Amazon India as it allows for price scraping.

```{r}
#the used Reference is "An introduction to web scraping using R"
# Link "https://www.freecodecamp.org/news/an-introduction-to-web-scraping-using-r-40284110c848/"

```

#### Loading the necessary packages
```{r}
library(xml2) #Used in web scraping
library(rvest) # "rvest is used in extracting the information we need from web pages."
library(stringr) #Used in data cleaning
```

#### Creating url variable + Read html content of the url
```{r}
url = "https://www.amazon.in/Samsung-Commercial-Displays-32-Inch-DC32E/dp/B01BJOF46K/ref=sr_1_3?keywords=samsung+tv+screen&qid=1637409358&s=electronics&sr=1-3"
webpage<- read_html(url)
```

#### Scrap the product details from the Amazon
#### The scraped data is as follows,
#### 1 - Title: The title of the product.
#### 2 - Price: The price of the product.
#### 3 - Description: The description of the product.
#### 4 - Rating: The user rating of the product."


```{r}
#scrape title of the product
title_html <- html_nodes(webpage, "h1#title")
title <- html_text(title_html)
head(title)
```
```{r}
#Removing /n and spaces,
title <- str_replace_all(title,"[\r\n]" , "")
title
```

#### Extracting other related information
#### A. Price
```{r}
#Price of the product
price_html <- html_nodes(webpage,"span#priceblock_ourprice")
price <- html_text(price_html)
price <- str_replace_all(price, "[\r\n]" , "")
price
```
#### B. Description
```{r}
desc_html <- html_nodes(webpage, "div#productDescription")
desc <- html_text(desc_html)
desc <- str_replace_all(desc, "[\r\n\t]" , "")
desc <- str_trim(desc)
head(desc)
```

#### C. Rating
```{r}
rate_html <- html_nodes(webpage, "span#acrPopover") 
rate <- html_text(rate_html)
rate <- str_replace_all(rate, "[\r\n]" , "")
rate <- str_trim(rate)
rate
```
#### Creating an R Data Frame

```{r}
#Creating a dataframe 
DataFrame<- data.frame("Title" = title, "Price" = price,"Description" = desc, "Rating" = rate)
DataFrame <- DataFrame[-2,] # Removing the dublicated row
DataFrame #Showing the data frame 

```

#### Data Frame metadata
```{r}
#Data Frame metadata
str(DataFrame)
```

#### Export to CSV file for further analysis
```{r}
write.csv(DataFrame,"/Users/salahkaf/Documents/Amazon_product.csv", row.names = FALSE)

```

