---
title: "Matrix Exercises"
author: "Salah"
date: "11/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R summation
```{r}
mat1 <- matrix(1:8, nrow = 4, ncol = 4)
mat2 <- matrix(data=1:16,nrow=4,ncol=4)
sum <- mat1 +mat2
sum
```
#In the same manner for subtraction, divisoin, multiplication
```{r}
diff <- mat1 - mat2
div <- mat1/mat2
Cprod <- mat1*4 #Constant Multiplication
prod <- mat1*mat2
MatMult <- mat1 %*% mat2
diff
div
Cprod
prod
MatMult
```

#Another example for clarification
```{r}

mata <- matrix(1:3,nrow=1)
matb <- matrix(1:6,nrow=3)
mata
matb
mata%*%matb
```


##### Write a R program to create three vectors a,b,c with 3 integers. Combine the three vectors to become a 3×3 matrix where each column represents a vector. Print the content of the matrix.
```{r}
a<-c(1,2,3)
b<-c(4,5,6)
c<-c(7,8,9)
row_names <-c ('row1','row2','row3')
column_names <- c('Vector a','Vector b','Vector c')
mat <- matrix(data=c(a,b,c),nrow=3,dimnames=list(row_names,column_names))
mat
```

##### R has a full suite of matrix-vector and matrix-matrix arithmetic operations, let
```{r}
x <- matrix (1:30 , 10) #Specifying ten rows
y <- matrix (1:10 , 2)
a <-1:3
b <-1:5
print(x+a) #Add the vector to the first row
print(x*x)
```
##### A square matrix A is said to be invertible if there exists a matrix B such that AB = BA = I, Where I is the identity matrix (1 along the diagonal, 0 elsewhere). 
```{r}
set.seed (1234)
x <- matrix( rnorm(25), nrow=5, ncol=5)
x
```

