---
title: "KNN with CV"
author: "Salah"
date: "1/5/2022"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Import the necessary libraries

```{r}
library('tidyverse') #For data frame manipulation and plotting
library('caret') #For machine learning
library('readxl') #For Excel reading
```

#### Read the Dataset and do descriptive analysis

```{r}
DF <- read_excel('/Users/salahkaf/Downloads/labW9.xlsx') #Read the excel file as a tibble
head(DF) #Shows top 6 rows
tail(DF) #Shows last 6 rows
dim(DF) #Shows number of columns and rows
str(DF) #Presents DF structure
summary(DF) #Presents DF summary
```

#### Checking for any missing values

```{r}
# Total number of missing values in the data set:
cat("The total number of missing values in the dataset is" , sum(is.na(DF)))
# Total number of missing values in the dataset per column name
colSums(is.na(DF)) 
```

#### Changing the "outcome" column (Target value) into categorical data to apply classification

```{r}
DF$Outcome<-gsub(1,"diabetic", as.character(DF$Outcome)) #Changing 1 to diabetic
DF$Outcome<-gsub(0,"Non-diabetic", as.character(DF$Outcome)) #Changing 0 to Non-diabetic
DF$Outcome <- as.factor(DF$Outcome) #Make it as a factor in order to apply classification
```


#### Checking the final form of the DF

```{r}
head(DF)
```

#### Splitting the dataset into 70/30 

```{r}
split = 0.7
trainIndex <- createDataPartition(DF$Outcome, p = split, list = F)
data_train <- DF[trainIndex,]
data_test <- DF[-trainIndex,]
```

#### Checking the training and testing subsets

```{r}
dim(data_train)
dim(data_test)
#### Equal number of columns (9), rows with 70% & 30% split
```

#### Apply cross validation

```{r}
train_control <- trainControl(method="cv", number=10) #10 subsets
```

#### Train the data using K nearest neighbor

```{r}
model <- train(Outcome~., data = data_train, trControl=train_control, method="knn")
```

#### Ploting the model

```{r}
plot(model)
```

#### Predict the values of test data

```{r}
predictions <- predict(model, newdata = data_test)
```

#Evaluate the outcome by using confusion Matrix

```{r}
cm <-confusionMatrix(predictions, data_test$Outcome)
cm
```

