---
title: "Data Frames Manipulation"
author: "Salah"
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.	Read file marks1.csv

```{r}
marks_df <- read.csv(file ='/Users/salahkaf/Downloads/marks1.csv')
marks_df
```
# 2.	Check the data frame info using a few available functions

```{r}
head(marks_df) #Top 6 rows
tail(marks_df) #last 6 rows
dim(marks_df) #Dimensions of DF
str(marks_df) #Structure of DF
summary(marks_df) #Summary of Df
```


# 3.	Check the names of the variables in the data frame

```{r}
colnames(marks_df) #Attributes names
```

# 4.	Rename the first variable X to ID
```{r}
colnames(marks_df)[1] <- "ID"
marks_df
```
# 5.	Rename the second variable X.1 to StuName
```{r}
colnames(marks_df)[2] <- "StuName"
marks_df
```
#6.	Remove the first two column from the data frame

```{r}
# First let us make a backup DF 
backup_df <- marks_df
# Removing column 1 & 2 
marks_df[,c(1,2)] <- NULL
marks_df
```

# 7.	Use apply() function to sum all the marks in the data frame and put them in a new vector called Total and bind the vector to the data frame

```{r}
Total <- apply(marks_df,1,sum)
Total
```

```{r}
marks_df$Total = Total
marks_df
```
# 8.	Using a user defined function called function(), use the apply() function to add variable 1 to variable 3, and write to a new variable in the data frame called CW. 

```{r}
cw_function <- function(x) {
  x[1]+x[2]+x[3]
}
cw <- cw_function(marks_df)
colnames(cw) <- "CW"
cw
```
```{r}
marks_df["CW"] <-cw
marks_df
```
# We stil need to restore the first two columns, i.e., ID and stuName from the backup_df
```{r}
ID <- backup_df[,1]
stuName <- backup_df[,2]
marks_df["ID"] = ID
marks_df["stuName"] = stuName
#Reordering the DF to ma
marks_df
```
# Relocating ID and stuName columns to the beginning 
```{r}
require(dplyr)
marks_df <- marks_df %>% relocate (ID, stuName,.before= test)
```
#Final DF
```{r}
marks_df
```

