---
title: "Data frames"
author: "Salah"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Data frame is a two dimensional data structure in R. It is a special case  of a list which has each component of equal length.
#### Data frame is a tabular data in R
#### Can store different classes of objects in each column
#### Data frames have a special attribute called row.names which indicates information about each row of the data frame.
#### Data frames can either by created by reading csv or tables file. read.table(..) or read.csv(..).
#### Or created explicitly using the data frame function

```{r}
data.frame( foo = 1:4 , bar = c(T, T, F, F)) 
```
#### A data frame is used for storing data tables. It is a list of vectors of equal length.
```{r}
n = c(2 , 3, 5)
s = c("aa", "bb", "cc") 
b = c(TRUE, FALSE, TRUE) 
df = data.frame(n, s, b)
df

```
#### Another example is as bellow,

```{r}
x <- data.frame("SN" = 1:2, "Age" = c(21,15),"Names" = c("Mohammad","Ahmed"))
x
str(x)
```
#### Basic functions for data frames,
```{r}
names(x)
nrow(x)
ncol(x)
length(x)
```
#### Accessing a column by its name
```{r}
x['Names'] 
x$Names 
class(x['Names']) #Column i.e., data frame
class(x$Names) #Vector of characters 
```
#### Accessing A certain element 
```{r}
# Both of them are the same output as follows,
x[["Names"]][2]
x$Names[2]

#Also,
x[1,] #Complete first row [df]
x[1:2,] #Complete first two rows [df]
x[,2] #Second row as a vector "Numeric" class(x[,2])
x[1,2] #first row second column value (Numeric) class(x[1,2])

```
#### To avoid creating x[,2] as a vector, we use the following code
```{r}
x1<- x[,2,drop = FALSE]
x1
class(x1)
#x["Age"] #This is dataframe
#x[["Age"]] #This is vector
```
#### Set new index names,
```{r}
row.names(x) <-c("stu1","stu2")
x
row.names(x) #Console output #Vector of characters
```

```{r}
names(x)[names(x)=='SN'] <- 'SID' #Renaming a column
x
names(x) <-c("StdID","StdAge","StdName") #Renaming all columns
x
```
#### Recap
```{r}

x[,3] #Character vector
x[,3,drop=FALSE] #DF
x["StdName"] #DF
x[["StdName"]] #Character vector
```

