---
title: "Tutorial 5 - EDA"
author: "Salah"
date: "12/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### attitude is the chosen dataset for this EDA

```{r}
attitude
```

#### 1.	See basic descriptive statistics

```{r}
dim(attitude) #Dimensions of Dataframe
str(attitude) #Structure of Dataframe
summary(attitude) #Summary of Dataframe
```

#### 1 - a)	What is the difference between (attitude[3]) and (attitude$learning)

#### Ans. 
#### Below, we observe that attitude[3] results a dataframe consists of one column. On the other hand attitude$learning produces a numeric vector

```{r}
attitude[3]
attitude$learning
class(attitude[3])
class(attitude$learning)

```

#### 2.	Lists name of variables in a dataset

```{r}
colnames(attitude)
```

#### 3.	Calculate number of rows in a dataset

```{r}
nrow(attitude)
```

#### 4.	Calculate number of columns in a dataset

```{r}
ncol(attitude)
```

#### 5.	List structure of a dataset

```{r}
str(attitude)
```

#### 6.	See first 6 rows (by default) of dataset

```{r}
head(attitude)
```

#### 7.	See first n rows of dataset. Select to see the first 15 rows of dataset

```{r}
head(attitude,15)
```
#### 8.	See all rows but the last row

```{r}
head(attitude,nrow(attitude)-1)
```


#### 9.	See last 6 rows (by default) of a dataset

```{r}
tail(attitude)
```


#### 10.	See last n rows of dataset. Select to see the last 12 rows of dataset.

```{r}
tail(attitude,12)
```
#### 11.	See the last n rows but the first row

```{r}
tail(attitude,nrow(attitude)-1) # See all rows but first row
```

#### 12.	Number of missing values. Which function will return number of missing values in each variable of a dataset?

```{r}
is.na(attitude)
cat("The total number of missing values in the dataset is" , sum(is.na(attitude)))
```

#### 13.	Number of missing values in a single variable

```{r}
for (i in 1:ncol(attitude)) {
x <- sum(is.na(attitude[i]))
cat("Number of missing values in col ", i, " = ", x, "\n")
}

```

#### 14.	Plot a simple graph, which will appear on a screen device.

```{r}
plot(attitude) #Using only the basic command
```


#### 15.	Plot the graph shown below, and make it appear on a file device (a pdf file)

```{r}
plot(x = attitude$privileges,
     y = attitude$learning,
     xlab = "Privileges",
     ylab = "Learning",
     main=  "Learning Attitude",
     xlim = c(30,85),
     ylim = c(40,75))

```

#### Saving the plot as PDF on my disktop 
```{r}
pdf(file = "/Users/salahkaf/Desktop/My Plot.pdf",   # The directory to save the plot
    width = 8, # The width of the plot in inches
    height = 8) # The height of the plot in inches
plot(x = attitude$privileges,
     y = attitude$learning,
     xlab = "Privileges",
     ylab = "Learning",
     main=  "Learning Attitude",
     xlim = c(30,85),
     ylim = c(40,75))
dev.off()
```

