---
title: "CodeBook Creation in R"
author: "Salah"
date: "11/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### This work is based on [Generate a simple codebook in R, https://www.adrianbruegger.com/post/simple-codebook/ ] accessed 21/11/2021.

#### First import the important haven library and a dataset 

```{r}
library(haven) # package to read files from popular statistical software packages such as SPSS, SAS, Stata
data <- read_sav("https://mmi.psycho.unibas.ch/r-toolbox/data/Cars.sav") # import data
```

#### Presenting the imported data (Data about cars)

```{r}
data
```
#### Showing the labels of variables using get_label and get_labels
```{r}
library(sjlabelled) # package to read and write item labels and values
get_label(data) # show content of variables (what the variable measures)

# which returns for example:
# MPG               
# "Miles per Gallon"

get_labels(data) # show value labels (what the different answer options mean)
# note: the value labels are not used for this very simple codebook.

# which returns for example:
# $CYLINDER
# [1] "3 Cylinders" "4 Cylinders" "5 Cylinders" "6 Cylinders" "8 Cylinders"
```

#### Create a simple codebook based on the labels of the dataset
```{r}
# extract labels from dataframe and store as new object
library(tibble)
simple_codebook <- enframe(get_label(data))

# use more informative column names
colnames(simple_codebook) <- c("variable_id", "item_text")

# Show the new data frame
simple_codebook
```

#### Generating decriptive statistics. 
```{r}
# get descriptive statistics and select those of interest
library(psych)
library(dplyr)
descriptives <- data %>% describe() %>% as_tibble() %>% select("n","min","max","mean")
# add stats to codebook 
simple_codebook <- cbind(simple_codebook,descriptives)
```


#### Saving the codebook in csv and xlsx file format. 

```{r}
# write to csv and Excel
write.csv(simple_codebook,file="simple_codebook.csv", na="", row.names=FALSE) 

library(openxlsx)
write.xlsx(simple_codebook,file="simple_codebook.xlsx")
```

