---
title: "Functions"
author: "Salah"
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Functions are implemented so we an reuse codes and make it easier to read them.
#### "Sharing methodology. y <- function(x1, x2, ...) “function” is the name of function to be run"
#### "y is the output of the function"
#### "x1, x2, . . . are the input arguments to the function"
#### " Functions are used to encapsulate a sequence of expressions that are  executed together to achieve a specific goal."

```{r}
#example of built-in functions
y <- mean(0:10) # or like this mean(x=0:10) = 5
z <- seq(20) #generate a sequal of consecutive 20 numbers [1,2,3,...,20]
```

### rnomr built-in function

```{r}
s <- rnorm(100) #Generate 100 random number
s <- rnorm(100) #Generate DIFFERENT 100 random number
s <- rnorm(10, mean =5, sd = 1.2) # mean = 5, Standard Deviation = 1.2
```

####  To fix the generated random numbers output

```{r}
### we the set.seed function when running simulations to ensure all results, figures, etc are reproducible.
set.seed(3)
rnorm(3, mean =5, sd = 1.2) # The three numbers will always be fixed
```

####Na family built-in functions

```{r}
z[3] <-NaN #Placing the third number is not a number NaN
is.na(z) # Checking each number if not available, output true or faluse # Index 3 is TRUE
anyNA(z)    # One output, is there a NA value? yes, then TRUE. 
```

#### Creating a data.frame containing a NA value.

```{r}
x=data.frame(a=c(2,3,5,8),b=c(3,8,NA,5),c=c(10,4,6,11))
x # Data frame contains a NA
anyNA(x) # TRUE
colMeans(x, na.rm = TRUE, dims = 1) #Removing the NA to complete the computation.
# a        b        c 
# 4.500000 5.333333 7.750000 
colMeans(x)
#   a    b    c 
# 4.50  NaN 7.75
```

#### Creating a function to calculate mean

```{r}
myMean <- function(x1) 
{
  return(sum(x1)/length(x1))
}
myMean(seq(1,10,1))
```

#### Defining a function to convert Fahrenheit to Celsius 
```{r}
FTC <- function(x) {
  Celsius <- (x-32)*(5)/9
  return(Celsius)
}
```

```{r}
FTC(312.52)
```

#### Generating a function to convet Celsius to fahrenheit

```{r}
CTF <- function(x) {
  Fahrenheit <- ((x*9)/5)+32
  return (Fahrenheit)
}
```

```{r}
CTF(200)
```
#### Nesting Functions is also possible
```{r}
STC <- function(x) {
  Kelvin <- x+273.15
  return(Kelvin)
}
STC(FTC(300)) #From 30 Fahrenheit to Kelvin
```

#### Overwriting functions

```{r}
myMean <-function(x) {
  cat("Function has changed")
}
myMean()
```


#### "R only allows return of a single object, but lists can be used for multiple outputs"

```{r}
myStats <- function (x) {
  m <- sum(x)/length(x)
  s<-sd(x)
  #return(c(MEAN=m,SD=s))
  #Output form
  #MEAN   SD 
  #10   NA 
  return(list(MEAN=m,SD=s))
  #Output form
  #$MEAN
#[1] 5.5

#$SD
#[1] 3.02765

#[1] 5.5
#[1] 3.02765
}
```
#### output of my myStats
```{r}
y <-myStats(1:10)
y #Return the complete list
y$MEAN #Return one element
y$SD #Return one element
```
#### Example pf multireturn 
```{r}
name_age_height<- function(name,age,height) {
  mylist <-list(NameofPerson=name,
              agefPerson=age,
              heightfPerson_cm= height)
  return(mylist)
}
```

```{r}
name_age_height("Arther",43,180)

```

#### "Write a function called edges that returns a vector made up of just the  first and last elements of its input?"

```{r}
elements <- c("element1","element2",
           "elemnt3","element4","element5")
edges<- function(x) {
return(list(first_element = x[1],last_element = tail(x,n = 1)))
}
```

```{r}
edges(elements)
```
#### "Write a function called top that returns takes a matrix or data frame, and returns the first n rows and columns :"
```{r}
#Create a random matrix
mat <- matrix(1:30,nrow=10,byrow=TRUE)
mat[1:5,]
```
```{r}
top5 <- function(x) {
  x[1:5,]
}
top5(mat)
```

#### Default values
#### "Functions can accept arguments explicitly assigned to a variable name  in the function call functionName(variable = value)."


```{r}
mySum <- function(input_1,input_2=10){
  output<- input_1+input_2
  return(output)
}
```

```{r}
mySum(3) #10+3
mySum(3,6) #3+6 will overwrite the value for input 2
mySum(input_2=9) #return error as there is no defalut value for input_1
```

#### Add messages throughout the function.

```{r}
myMean <- function(x) {
  if(any(is.na(x))) {
    message("x shoul not contain NA values")
  }
  s<- sum(x)
  message(paste("the sum of x is", s))
  l<- length(x)
    message(paste("the length of x is", l))
    message("The mean is")
    return(s/l)
}
```

```{r}
myMean(1:3)
```

