---
title: "Grouping, loops & conditional execution"
author: "Salah"
date: "11/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Grouped expressions:
#####  "R is an expression language in the sense that its only command type is a function or expression which returns a result. Commands may be grouped together in braces, {expr 1, . . ., expr m}, in which case the value of the group is the result of the last expression in the group evaluated."

#### Control Statement 
##### "if statements –The language has available a conditional construction of the form if (expr 1) expr 2 else expr 3 where expr 1 must evaluate to a logical value and the result of the entire expression is then evident. a vectorized version of the if/else construct, the ifelse  function. This has the form ifelse(condition, a, b)"

#### Repetitive execution as  For loops, repeat and while
##### Branching is optional 
##### R is not indent sensetive, but indent are preferably to do in conditions and {} is a must.

```{r}
voter <-T
age <-10
if (age <18) {
  message("This is a minor")
  voter <- F
}
message("If statement is completed")
voter
```

```{r}
voter <-T
age <-20
if (age <18) {
  message("This is a minor")
  voter <- F}

message("If statement is completed")
voter
```
##### Another concise way is as follows,
```{r}
age<-18
if (age >=18) {
  message("This is an adult")
  voter <- T
} else {
  message("This is a minor")
  voter <- F
}
voter
```
##### IFELSE statement quick assigment for a vector
```{r}
ages <- c(12,18,32,2,4)
names(ages) <- c("Jane","Ed","Bob","Joe","Liz")
voter <- ifelse(ages<18,'minor','voter') #minor is condition for IF, voter is condition for else
voter
```
#If else example
```{r}
age <-20
if (age <18) {
  message("This is a minor")
  voter <-F
  driver<-F
} else if (age <21) {
  message("This is a voter")
  voter <-T
  driver<-F
} else {
  message("This is an adult")
  voter <-T
  driver<-T
}
voter
driver

```

#### switch command works as ifelse
```{r}
dish <- "fruit"
switch(dish, fruit ="banana",vegetable ="broccoli", meat="chicken") #Which option did the customer choose? it is fruit!
#In the same manner, here we go with order number 3!
x<-switch(3,'first','second','third','fourth')
x
```

#### Loops: "When the same or similar tasks need to be performed multiple times; for all elements of a list; for all columns of an array; etc."

```{r}
for (i in 1:10) {
  print(i*i)
}
```
#### Another example,
```{r}
i = 1
while (i<=10) {
  print (i*i)
  i= i+1
}
```

##### " For loops repeat the same task for changing values of a given variable."

```{r}
ages <-c(12,18,32,2,4)
names(ages) <- c("Jane","Ed","Bob","Joe","Liz")
for (person in names(ages)) {
  message(person, " is ", ages[person], " years old")
}
```

```{r}
agesSum <-0
ages<-c(12,18,32,2,4)
names(ages) <-c('Jane','Ed','Bob','Joe','Liz')
for (person in names(ages))
{
  message(person, ' is ', ages[person], ' years old.')
  agesSum <-agesSum+ages[person]
}
agesSum
```

#### Nested loop # A loop inside a loop

```{r}
ages<-c(12,18,32,2,4)
names(ages) <-c('Jane','Ed','Bob','Joe','Liz')
for (person in names(ages)) {
  message(person)
  for(i in 1:ages[person])
  {
    message("Year",i)
  }
}
  
```
#### Adding if statement in for loops
```{r}
ages <- c(12,18,32,2,4)
numVoters <-0
names(ages) <-c('Jane','Ed','Bob','Joe','Liz')
for(person in names(ages))
{
  if (ages[person]>=18)
  {
    message(person," can vote")
    numVoters <- numVoters+1
  }
}
numVoters
```
#### Setting values with embedded if statement in for loops

```{r}
ages <- c(12,18,32,2,4,40)
numVoters <-0
names(ages) <-c('Jane','Ed','Bob','Joe','Liz','steve')
voters <-vector()
for (person in names(ages))
{
  if (ages[person]>=18)
  {
    message(person, "can vote")
    numVoters <-numVoters +1
    voters[numVoters] <- person #numVoters is the index for voters vector
  }
}
numVoters
voters
```

#### While loop
##### while loops repeat the same operation until a test condition is met.

```{r}

z <-0
while(z<5)
{
  z<- z+1
  message(z)
}
```

#### Repeat loop and break statement
##### "A break statement is used inside a loop (repeat, for, while) to stop the iterations and flow the control outside of the loop."

```{r}
x <- 1
repeat {
  print(x)
  x = x+1
  if (x==6){
    break
  }
}
```


```{r}
x <- 1:5
for (val in x) {
  if (val==3) {
    break #Stop at 3
}
print(val)
}
```

```{r}
x <- 1:5
for (val in x) {
  if (val==3) {
    next #skip 3
}
print(val)
}
```

