---
title: "Subsetting"
author: "Salah"
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### subsetting is selecting (and keeping) or deleting (dropping) variables and observations, or to take random samples from a dataset.

```{r}

#regular subsetting 

x<- data.frame("SN"=1:4,
          "Age"=c(21,20,25,27),
          "Name"=c("John","Dora","Mary","Jim"), "Height" = c(162,157,155,129), "Gender" = c("M","F","F","M"))
x[c("Age","Name")] #include age and name
x[c(-2,-3)] #Removing second and third row
```
```{r}
x$SN <- x$Age <- NULL #delete rows permanently
```

```{r}
x
```

```{r}
hsb2.small <- read.csv("https://stats.idre.ucla.edu/stat/data/hsb2_small.csv")
```

```{r}
names(hsb2.small) #Reading the columns names
```

```{r}
# Regular subsetting
hsb3 <- hsb2.small[,c(1,7,8)] #Selecting specific columns
hsb3
```
```{r}
hsb4 <- hsb2.small[, 1:4] #colon notation
hsb4
```
```{r}
hsb6 <- hsb2.small[hsb2.small$ses==1,] # selecting the rows where ses =1 
hsb6
```
```{r}
hsb7 <-hsb2.small[hsb2.small$id %in% c(12, 48, 86, 11, 20, 195),] #The condition here is id value must exist in the instructed vector %in%
hsb7
```
```{r}
hsb8 <- hsb2.small[hsb2.small$ses==3 & hsb2.small$female==0,]
hsb8
#Another way is using with to not repeat hsb2.small
hsb8b <- hsb2.small[with(hsb2.small,ses==3,female==0),]
if (hsb8==hsb8b) {
  print("Match!")
}
```
#### Subsetting is an easier way to select variables and observations.
#Subsetting using subset function as below,


```{r}

mydata<- data.frame("SN"=1:4,
          "Age"=c(21,20,25,27),
          "Name"=c("John","Dora","Mary","Jim"), "Height" = c(162,157,155,129), "Gender" = c("M","F","F","M"))
mydata
```


```{r}
# selecting the original set, then the conditions, and which columns
newdata <- subset(mydata, Age>22 | Age <20,select=c(SN,Name))
newdata
```
```{r}
#Another example of subsetting
newdata2 <- subset(mydata,Gender =="M" & Age > 25, select= SN:Height) #Select from column SN to Height column
newdata2
```

#### The subset function with a logical statement will let you subset the data frame by observations. 

```{r}
names(hsb2.small)
hsb2.small
write.50 <- subset(hsb2.small, write > 50)
write.50
write.1 <- subset(hsb2.small,write>50 & read >60)
write.1
```
#### Subsetting rows and columns together
```{r}
write.2 <- subset(hsb2.small,write>50 & read>60, select=c(write,read))
write.2

#In the data frame write.3 contains only the observations in variables read through science for which the values in the variable science are less than 55.
write.3 <- subset(hsb2.small, science < 55, select = read:science)
write.3
```

