---
title: "PLotting Basics"
author: "Salah"
date: "1/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Base R plots

```{r}
# plot(x, y, type="p")
# x and y: the coordinates of points to plot
# type : the type of graph to create; Possible values are :
# type=“p”: for points (by default)
# type=“l”: for lines
# type=“b”: for both; points are connected by a line
# type=“o”: for both ‘overplotted’;
# type=“h”: for ‘histogram’ like vertical lines
# type=“s”: for stair steps
# type=“n”: for no plotting
```

```{r}
data("mtcars")
plot(mtcars) #Overall plot between attributes
str(mtcars) 
```

#### Basic R plots
```{r}
plot( mtcars$wt, mtcars$mpg) #default type p #MPG  Y axis
plot(mtcars$wt, mtcars$mpg, type = "l")
plot(mtcars$wt, mtcars$mpg, type = "p")
plot(mpg ~ wt, data = mtcars) #MPG Y axis
plot(wt ~ mpg, data = mtcars) #WT is the dependent variable - Y axis
# plot(wt, mpg, data= mtcars) #error

plot(mtcars$wt, mtcars$mpg, main = "mtcars dataset", xlab = "X axis wt", ylab = "Y axis mpg")

plot(mtcars$wt, mtcars$mpg, main = "mtcars dataset",  xlab = "X axis wt", ylab = "Y axis mpg",
     pch = 19, frame = FALSE)
lines(lowess(mtcars$wt, mtcars$mpg), col = "blue")

```

#### ggplots syntax

#### All the data needed to make the plot is typically be contained within  the dataframe supplied to the ggplot() itself or can be supplied to  respective geoms.
#### Another noticeable feature is that you can keep enhancing the plot by  adding more layers (and themes) to an existing plot created using the  ggplot() function

```{r}
# library('tidyverse')
# qplot in ggplot2 package
qplot(wt, mpg, data = mtcars, geom= c("point"))
qplot(wt, mpg, data = mtcars, geom= c("smooth"))
qplot(wt, mpg, data = mtcars, geom= c("point","smooth"))


qplot(wt, mpg, data = mtcars, color=cyl)  #better if char or factor

qplot(mpg, wt, data = mtcars, colour = cyl) #inverted axis 

qplot(mpg, data = mtcars, fill=cyl )  #the data doesnt reflect

# Three variables in 2 D plot (One is categorical)
mt<-mtcars
mt$cyl <- as.factor(mt$cyl)
qplot(mpg, wt, data = mt, colour = cyl)  #better with factor data

```

#### Understanding ggplot

#A blank ggplot is drawn. Even though the x and y are specified, there  are no points or lines in it. This is because, ggplot doesn’t assume  that you meant a scatterplot or a line chart to be drawn.
#Also note that aes() function is used to specify the X and Y axes.  Any information part of the source dataframe has to be specified  inside the aes() function

```{r}
ggplot(mtcars, aes(wt, mpg)) + geom_point()

ggplot(mtcars, aes(wt, mpg)) + geom_point() + geom_smooth() #default method

ggplot(mtcars, aes(wt, mpg)) + geom_point() + geom_smooth(method = "lm")

ggplot(mtcars, aes(wt, mpg)) + geom_point(aes(color=cyl))

# Three variables in 2 D plot (One is categorical)
ggplot(mtcars, aes(wt, mpg)) + geom_point(aes(color= factor(cyl))) #can change to factor

ggplot(mtcars, aes(x = factor(cyl))) +  geom_bar()

```

#### Add labels and titles

```{r}
g2<-ggplot(mtcars, aes(wt, mpg)) + geom_point(aes(color= factor(cyl))) #can change to factor
g2 + labs( title =" Mpg vs Wt ", subtitle =" optional title", 
           y=" mpg ", x=" wt ", caption =" mtcars Dataset ")

```
```{r}
# #or
# g2  +  ggtitle (" Mpg vs Wt ", subtitle =" optional title") + 
#   xlab (" Mpg ") + ylab (" wt ")
# #or full
# ggplot(mtcars, aes(wt, mpg)) + geom_point(aes(color= factor(cyl))) + 
#   labs( title =" Mpg vs Wt ", subtitle =" optional title", 
#         y=" mpg ", x=" wt ", caption =" mtcars Dataset ")

```

