---
title: "Dplyr"
author: "Salah"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Subsetting with base R is achived via index or names.
# Subsetting wihch() and with()

```{r}
 # All columns, rows only cyl = 8
mtcars[mtcars$cyl==8,]
mtcars[which(mtcars$cyl==8),]
mtcars[with(mtcars,cyl==8),]
# All of them are data frames
```

```{r}
#Row condition cyl = 8, column number 1 
mtcars[mtcars$cyl == 8, 1]
mtcars[which(mtcars$cyl == 8), 1]
mtcars[with(mtcars, cyl == 8), 1]
#Class numeric
```

```{r}
#Row condition cyl = 8, column number 1:3
mtcars[mtcars$cyl == 8, 1:3]
mtcars[which(mtcars$cyl == 8), 1:3]
mtcars[with(mtcars, cyl == 8), 1:3]
#class data frames

```

# More specific subsetting
```{r}
#Two row condition cyl=8 and gear =3 
mtcars[mtcars$cyl==8 & mtcars$gear>3,1:3]
mtcars[which(mtcars$cyl==8 & mtcars$gear>3),1:3]
mtcars[with(mtcars,cyl==8 & gear>3),1:3]
```

#Subsetting using subset (Row condition)

```{r}
#If columns are not specified, it will take all by default
subset(mtcars,cyl==8)
subset(mtcars,cyl==8 & gear ==5)
subset(mtcars,cyl==8 & gear < 5)

#Specifying the columns 
subset (mtcars, cyl == 8, select= c(1:3))
subset (mtcars, cyl == 8, select= c(1:3, 7))

```

```{r}
#Adding select within subset provides column condition
#Subset it based on row conditions
subset (mtcars, cyl == 8 & gear< 5, select = c(1:3, 10))
subset(mtcars,cyl==8 &gear <5, select=mpg:disp) #From column mpg to disp
```


# Dplyr family functions

# glimpse function
# "One handy feature with dplyr is the glimpse(..) function. This is  an improvement over str(..)."
# "Use glimpse(..) to see the structure of the dataset and  decide  what manipulation is required."

```{r}
library(dplyr)
glimpse(mtcars)
```

# sum () counts the observations in a column in the dataframe
```{r}
sum(mtcars$cyl==8) #Number of cars which have 8 cyl in the data frame
sum(mtcars$gear>4) #Number of cars which have a gear > 4 
```

# In dplyr package, 
# select() picks variables based on their names.
# mutate() adds new variables that are functions of existing variables
# filter() picks cases based on their values.
# arrange() changes the ordering of the rows.
# summarise() reduces multiple values down to a single summary. 
# Pipeline

```{r}
select(mtcars,1:3)
select(mtcars,-7) #All but 7th column
select(mtcars,cyl:wt)
select(mtcars, cyl:wt, -hp)
select(mtcars, 1:4, -3)
```
# Via helper functions in select, we can even be more specific in our indexing as bellow,
```{r}
select (mtcars, starts_with("c"))
select (mtcars, starts_with("c"),   ends_with("t"))
select (mtcars, matches("car"),matches("dra"))

# Dplyr has helper functions to help select a group of variables inside a select()
# starts_with(“x”) 	every name that starts with “x”
# ends_with(“x”)	every name that ends with “x”
# contains(“x”)	every name that contains “x”
# matches(“x”)	every name that matches “x”, where “x” can be a regular expression
# num_range(“x”, 1:5)	the variable named x01, x02, x03, x04, x05
# one_of(x)		every name that appears in x, which should be a character vector
# 
# Note : when referring to variables directly inside select(), don’t use quotes “”. When using helper function, use quotes

```

# Mutate() adds new column based on previous ones
```{r}
mutate(mtcars,hpp=hp/100)
# Creates new column and add to a copy of the dataset
# Syntax   
# mutate(df, z=x+y)
# Can add multiple variables
# Can also use the newly created variable
# mutate(df, z=x+y, zz = z+v)

```

# The filter(..) works exactly like select(), + you pass the data  frame first and then a condition separated by a comma

```{r}
filter(mtcars,cyl==8, qsec>16)
filter(mtcars,cyl==8 & qsec>16)

#Checking the number of rows using sum 

filter(mtcars,cyl==8,qsec>16) #10 rows
sum(mtcars$cyl==8 & mtcars$qsec >16) #10 rows
filter(mtcars,(cyl==8 | cyl==6) & qsec >16) #16 rows
sum( (mtcars$cyl==8 |mtcars$cyl==6) & mtcars$qsec>16) #16 rows
```

# Filter() Can use logical operators inside filter()
# < , <= , == , != , >= , >
# %in%


```{r}
filter(mtcars,cyl %in% c(4,8)) #No. of cyl exist in the vector (4,8)
filter(mtcars,is.na(x)) # is.na()	for observations x where the value is NA

```

#Arrange ()

```{r}
arrange(mtcars,cyl) #based on number cyl
arrange(mtcars, cyl, mpg) #Cyl first, then mpg
arrange(mtcars,cyl,desc(mpg))
```

#Arranging a filtered data set
```{r}
arrange(filter(mtcars,(cyl==8 | cyl ==6) & qsec > 16), desc(cyl))
```
# Arranging a filtered and selected df

```{r}
arrange(filter(select(mtcars,mpg:vs,-disp,-drat),(cyl==8 | cyl ==6) & qsec >16), desc(cyl))
```
# "summarise() creates a new data frame. It will have one (or more) rows for each combination of grouping variables; if there are no grouping variables, the output will have a single row summarising all observations in the input. It will contain one column for each grouping variable and one column for each of the summary statistics that you have specified."

```{r}
summarise(mtcars, min_mpg=min(mpg))
summarise(mtcars, maximum_mpg=max(mpg),min_mpg=min(mpg))
summarise(mtcars, mean = mean(mpg), n = n())
```

# "In summarise(), we can use any aggregating function that can take a vector of data and return a single number including "min(), max(), mean(), median(), quantile(), sd(), var()""

```{r}
summarise(filter(mtcars, cyl ==8), min_mpg = min(mpg), max_mpg = max(mpg), sd = sd(mpg), var = var(mpg))
```
# The dplyr library comes with a practical operator, %>%, called the  pipeline.
# The pipeline feature makes the manipulation clean, fast and less  prompt to error.
# define  the data frame used at the beginning and all the process will flow  from it.

```{r}
#Original Code,

arrange(filter(select (mtcars, mpg:vs, -disp, -drat), (cyl == 8 | cyl == 6) & qsec >16), desc(cyl))
#Pipeline below,

newmtpipe <- mtcars %>%
  select(mpg:vs,-disp,-drat) %>%
  filter((cyl==8 | cyl ==6) & qsec >16) %>%
  arrange(desc(cyl))
newmtpipe
```
#Another example of pipeline

```{r}
#Original code,
summarise(filter(mtcars,cyl==8), min_mpg=min(mpg), max_mpg = max(mpg))
#Pipeline
sumpipe <- mtcars %>%
  filter(cyl==8) %>%
  summarise(min_mpg = min(mpg), max_mpg=max(mpg))
sumpipe
```

# Exercise
## Load the iris dataset
```{r}
iris
```
## Select the first three columns of the iris dataset using  their column names. HINT:Use select().

```{r}
select(iris,Sepal.Length,Sepal.Width,Petal.Length)
```
## Select all the columns of the iris dataset except "Petal Width". HINT: Use -.
```{r}
select(iris,-Petal.Width)
```
## Filter the rows of the iris dataset for Sepal.Length >=  4.6 and Petal.Width >= 0.5.

```{r}
filter(iris, Sepal.Length >= 4.6 & Petal.Width >= 0.5)
```
## Create a new column called proportion, which is the ratio of Sepal.Length to Sepal.Width. HINT: Use mutate().

```{r}
mutate(iris, proportion = Sepal.Length/Sepal.Width)
```

