---
title: "Missing Values"
author: "Salah"
date: "12/21/2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Use is.na() to find NA values in dataset
####  is.na() returns a logical vector with TRUE in the element locations that contain missing values represented by NA. is.na() will work on vectors, lists, matrices, and data frames.

```{r}
x <- c(1,2,3,NA)
is.na(x) #Return a vector
anyNA(x) #Returns a boolean value
#  Use !is.na() to negate that
!is.na(x)
```

####Exlude missing value from analysis 
```{r}
mean(x) # = NA
mean(x,na.rm=TRUE) #6/3
```

####Setting a particulat data to NA

```{r}
ID <- 1:4
Gender <- c('M','F','','M')
Names <- c('', "Sarah", "Semanth", "David")
DF <- data.frame(ID,Gender,Names)
DF
```

```{r}
DF$Names[DF$Names==''] <-NA
DF$Gender[DF$Gender==''] <-NA
DF
```
#### Compelete.cases checks each row for NA value
#### False means there is a missing value
```{r}
complete.cases(DF) # Checking by row
```

```{r}
is.na(DF) # Shows in each row where the NA is
anyNA(DF) #Is there an NA?
any(is.na(DF))  #Is there an NA?
summary(DF)
```

#Removing rows with NA values
```{r}
DF
Cleaned_Df <- na.omit(DF)
Cleaned_Df
```

#### Test for missing values

```{r}
x <- c(1:4, NA, 6:7, NA)
x
is.na(x)
```

#### DF with missing values
```{r}
df <- data.frame(col1 = c(1:3, NA),
                 col2 = c("this", NA,"is", "text"), 
                 col3 = c(TRUE, FALSE, TRUE, TRUE), 
                 col4 = c(2.5, 4.2, 3.2, NA),
                 stringsAsFactors = FALSE)
df
```

```{r}
is.na(df) #Whole DF
is.na(df$col2) #Certain column
```

#### To identify the location or the number of NAs we can leverage the which() and sum() functions:

```{r}
which(is.na(x)) # # identify location of NAs in vector
sum(is.na(df)) # identify count of NAs in data frame
colSums(is.na(df)) # For data frames, a convenient shortcut to compute the total missing values in each column is to use 
```


#### Recode missing values

```{r}
# recode missing values with the mean
# vector with missing data
x <- c(1:4, NA, 6:7, NA)
x

```

```{r}
x[is.na(x)] <- mean(x, na.rm = TRUE) #Replacing NA values with the mean
x
```

```{r}
round(x, 2) #Round to two digits 
```

```{r}
# data frame that codes missing values as 99
df <- data.frame(col1 = c(1:3, 99), col2 = c(2.5, 4.2, 99, 3.2))
df
```

```{r}
df[df == 99] <- NA #Replacing 99 with NA
df

```

#### For finding the mean of one column and replacing NA 

```{r}
df <- data.frame(col1 = c(1:3, NA),
                 col2 = c("this", NA,"is", "text"), 
                 col3 = c(TRUE, FALSE, TRUE, TRUE), 
                 col4 = c(2.5, 4.2, 3.2, NA),
                 stringsAsFactors = FALSE)
df
```

#### Finding the NA values in a specific column, then replacing those NAs with the mean of that column
```{r}
df$col4[is.na(df$col4)] <- mean(df$col4, na.rm = TRUE)
df
```



#### Exclude missing values

```{r}
x <- c(1:4, NA, 6:7, NA)
mean(x)
```

```{r}
mean(x, na.rm = TRUE)
```


```{r}
df <- data.frame(col1 = c(1:3, NA),
                 col2 = c("this", NA,"is", "text"), 
                 col3 = c(TRUE, FALSE, TRUE, TRUE), 
                 col4 = c(2.5, 4.2, 3.2, NA),
                 stringsAsFactors = FALSE)

df

```

```{r}
complete.cases(df)
```

```{r}
# subset with complete.cases to get complete cases
df[complete.cases(df), ]
##   col1 col2 col3 col4
## 1    1 this TRUE  2.5
## 3    3   is TRUE  3.2

# or subset with `!` operator to get incomplete cases
df[!complete.cases(df), ]
##   col1 col2  col3 col4
## 2    2 <NA> FALSE  4.2
## 4   NA text  TRUE   NA
```

#### If we want to exclude missing values from mathematical operations use the na.rm = TRUE argument.


```{r}
# or use na.omit() to get same as above
na.omit(df)
##   col1 col2 col3 col4
## 1    1 this TRUE  2.5
## 3    3   is TRUE  3.2

```

#### Detecting outlies via histogram or boxplot are good approaches
