---
title: "TIDYR"
author: "Salah"
date: "12/8/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### "Whenever working with rectangular data structures — data consisting of multiple cases (rows) and variables (columns) — our first step (in a tidyverse context) is to create or transform the data into a tibble. A tibble is a rectangular data table and a modern and simpler version of the data.frame construct in R."

```{r}
library(tidyverse)
```

#### "Whereas data frames were present in base R in the 1990s (and a part of the S programming language on which R is based), tibbles first appeared in the dplyr package and in the form of an R package tibble (v1.0) in 2016. Nevertheless, tibbles essentially are simpler data frames. In contrast to the base R behavior of data frames, turning data into tibbles is more restrictive. Specifically,"
#### "tibbles do not change the types of input variables (e.g., strings are not converted to factors);"
#### "tibbles do not change the names of input variables and do not use row names."
#### "Tibbles are quite flexible in also allowing for non-syntactic variable (column) names. For instance, in contrast to data frames, the variable names in tibbles can start with a number or contain spaces"


```{r}
mtc <- as_tibble(mtcars)
class(mtc)
mtc
```

# TIDYR FAMILY 

### gather() collapse multiple columns into two columns
### takes multiple columns, and gathers them into key-value pairs
### Convert wide to long data
### gather(df, key, value, column)

```{r}
# Creating a messy df
messy_df <- data.frame(name=c('Wilbur','Petunia','Gregory'),
                       a=c(67,80,64),
                       b=c(56,90,50))
messy_df
```
# Applying Gather
```{r}
## gather(df, key, value, column) #Key takes the columns names # Value take the numbers in rows 
tidy_df <- gather(messy_df,drug,heartrate,2:3)
tidy_df
```
## Spread() converts long data to wide
## takes two columns (key&value) and spreads into multiple columns
# The opposite of gather

```{r}
messy_df1 <- spread(tidy_df,drug,heartrate)
messy_df1
# In other words spread is the opposite of gather
```

## Another spread example 

```{r}
original_df <- data.frame(
country=c("Afghanistan","Afghanistan","Afghanistan","Afghanistan","Brazil","Brazil","Brazil","Brazil"),
year = c(1999,1999,2000,2000,1999,1999,2000,2000),
key =c("cases","population","cases","population","cases","population","cases","population"),
value =c(745,19987071,2666,20595360,37737,172006362,80488,174504898))
original_df

```

```{r}
spread_original_df <- spread(original_df,key,value)
#Keys become columns, in other words, variables in one column goes into columns
#Values are spreaded over rows
spread_original_df
```

## Another example of spread
```{r}
row <- c(1,1,1,51,51,51)
vat <- c('Sepal.Length','Species','Species_num','Sepal.Length','Species','Species_num')
value <- c(5.1,'setosa',1,7,'versicolor',2)
df <- data.frame(row,vat,value)
df
```
```{r}
spread_df <- spread(df,vat,value)
spread_df
```

```{r}
city <- c("New York","New York","London","London","Beijing","Beijing")
size <- c("large","Small","large","Small","large","Small")
amount <- c(23,14,22,16,121,56)
df <- data.frame(city,size,amount)
df
```

```{r}
spread_df <- spread(df,size,amount)
spread_df
```

## Separate()
## Multiple variables are stored in one column are then separated into multiple columns
## Separete(df,data,colname)

```{r}
# Creating a messydf
messydf <- data.frame(
  name = c('Jake','Alice','Tim','Denise'),
  sex_age = c("M.34",'F.55','M.76','F.19'),
  eye_color = c("Other",'Blue','Brown','Other'),
  Height = c('6\'1"','5\'9"','5\'7"','5\'1"'))
messydf
```

## Applying separation

```{r}
tidydf <- separate(messydf,2,c("sex","age"))
tidydf
```

## Another example of separation

```{r}
  x <- c(NA,'a.b','a.d','b.c')
  y <- c(12,12,13,14)
  df <- data.frame(x,y)
  df
```

```{r}
separate(df,x,c("colA","colB"))
```

## Unite()
## To join variables
## unite(df, newcol, original_columns)
```{r}
head(mtcars)
```

```{r}
unitedmtcars<- unite_(mtcars,"vs_am",c("vs","am"))
head(unitedmtcars)
```

#example of messy data handling

```{r}
id <- 1:4
trt <- c("treatment","control","treatment","control")
work.T1 <-c(0.0851,0.2254,0.2745,0.2723)
home.T1 <-c(0.6158,0.4296,0.6516,0.5677)
work.T2 <-c(0.1135,0.5959,0.3580,0.4288)
home.T2 <-c(0.5190,0.2641,0.3987,0.8361)
messy <- data.frame(id,trt,work.T1,home.T1, work.T2,home.T2)
messy
```

#tidier below but still messy

```{r}
tidier <- gather(messy,key,time,3:6)
tidier
```
## Work.T shoulb be separate two columns, one is for location, and one is for time

```{r}
# tidy <- separate(tidier,key,c("Location","Time"))
# tidy
tidy <- tidier %>%
  separate(key,into=c("Location","Time"),sep="\\.")
tidy
```

## tidy data can still be cleaned. Time column has two values that can be made into 2 variables, T1 and T2

```{r}
even_tidier <- tidy %>%
  spread(Time,time)
even_tidier
```

